---
title: "Project Course - Pratical Machine Learning"
author: "Rodrigo Alves dos Santos"
date: "7 de novembro de 2016"
output: word_document
---

## Prediction Assignment Writeup

I - Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 
More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

Full Source: Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. "Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13)". Stuttgart, Germany: ACM SIGCHI, 2013.

A short description of the datasets content from the authors':
"Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).
Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes. Participants were supervised by an experienced weight lifter to make sure the execution complied to the manner they were supposed to simulate. The exercises were performed by six male participants aged between 20-28 years, with little weight lifting experience. We made sure that all participants could easily simulate the mistakes in a safe and controlled manner by using a relatively light dumbbell (1.25kg)."

We will use two diffente methods to classifie the participants and chose a method with a bigger accurary

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```{r, include=TRUE}
## Course Project - Pratical Machine Learning
# Rodrigo Alves dos Santos

# We start uploading the libraries that are necessary

library(corrplot)
library(caret)
library(Hmisc)
library(knitr)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(RColorBrewer)

knitr::opts_chunk$set(echo = FALSE)
# Setting the seed

set.seed(12345)

# We start downloading the data from URL

UrlTrain <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
UrlTest  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

# We wil use the testing set in URL as a Validation Set.
training <- read.csv(url(UrlTrain))
validation  <- read.csv(url(UrlTest))

# Exploratory Analysis
dim(training)

# create a partition with the training dataset 
inTrain  <- createDataPartition(training$classe, p=0.6, list=FALSE)
Train_Set <- training[inTrain, ]
Test_Set  <- training[-inTrain, ]

# Exploratory Analysis on training and testing set
dim(Train_Set)
dim(Test_Set)

# The dataset has 160 variables. 
# We will remove, with the cleaning procedures, the Near Zero variance (NZV) variables and the Identification (ID) variables.

# remove ID variables (columns 1 to 5)
Train_Set <- Train_Set[, -(1:5)]
Test_Set  <- Test_Set[, -(1:5)]

dim(Train_Set)
dim(Test_Set)

# remove variables NZV.
NZV <- nearZeroVar(Train_Set)
Train_Set <- Train_Set[, -NZV]
Test_Set  <- Test_Set[, -NZV]

dim(Train_Set)
dim(Test_Set)

# remove variables that are mostly NA.
MNA <- sapply(Train_Set, function(x) mean(is.na(x))) > 0.90
Train_Set <- Train_Set[, MNA==FALSE]
Test_Set  <- Test_Set[, MNA==FALSE]

dim(Train_Set)
dim(Test_Set)

# Now, the number of variables has been reduced to 54.


```

A correlation analysis is very important in a regression problem, but is less important in classfication problem. Anyway, we will look on the Correlation Graph.

```{r, echo=TRUE}
# Correlation Analysis

corMatrix <- cor(Train_Set[, -54])
corrplot(corMatrix, order = "FPC", method = "color", type = "lower", 
         tl.cex = 0.8, tl.col = rgb(0, 0, 0))

```

There is some correlated variables, but there is acceptable
Principal Components Analysis could be used as pre-processing step to reduce dimensionality, but, the correlations are quite few, this step will not be applied, because we dont wanna losse interpretability of the variables

```{r, echo=TRUE}
## model fit

# CrossValidation with 10 folds will be used for robust results
controlRF <- trainControl(method="cv", number=10, verboseIter=FALSE)

#Decision Tree
DecisionTree <- rpart(classe ~ ., data=Train_Set, method="class")
fancyRpartPlot(DecisionTree)

#Random Forest
RandForest <- train(classe ~ ., data=Train_Set, method="rf",
                          trControl=controlRF)
RandForest$finalModel


## Comparing the results in Testset (Test_Set)

# Random Forest
predictRF <- predict(RandForest, newdata=Test_Set)
confMatRF <- confusionMatrix(predictRF, Test_Set$classe)
confMatRF

plot(confMatRF$table, col = confMatRF$byClass, 
     main = paste("Random Forest - Accuracy =",
                  round(confMatRF$overall['Accuracy'], 4)))

# Decision Tree
predictDecisionTree <- predict(DecisionTree, newdata=TestSet, type="class")
confMatDecisionTree <- confusionMatrix(predictDecisionTree, TestSet$classe)
confMatDecisionTree

plot(confMatDecisionTree$table, col = confMatDecisionTree$byClass, 
     main = paste("Decision Tree - Accuracy =",
                  round(confMatDecisionTree$overall['Accuracy'], 4)))
```

Random Forest is the method with higher accuracy, them we will applied him in the Validation Set

```{r, echo=TRUE}
# We use Random Forest in a Validation Set, the 20 cases that is provide as a Testing Set.

predict_Validation <- predict(RandForest, newdata=validation)
predict_Validation
```